---
title: "DA5020.A1.NOAH.DCRUZ"
output:
  html_document: default
  word_document: default
  pdf_document: default
date: "2024-01-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Question 1
### A dataframe is a table like structure that has columns(variables) and rows(data points or observations)
```{r}
s <- data.frame(ID = 1:3, Name = c("N1","N2","N3"),height=c(60,70,80))

s
```

### Question 2

```{r}
cars <- c("Truck","Car", "SUV")
mpg <- c(11, 30, 24)
cost <- c(45000,25000,35000)

df <- data.frame(cars, mpg, cost)

print(df)
```

### Question 2a

```{r}
df[1,3]
df[1:3,]
df[,ncol(df)]
```

### Question 3

```{r}
head(mtcars,3)
tail(mtcars, 5)
```


### Question 4
## Firdt one is categorical and second is continuous
```{r}
str(mtcars)
cat <- c("vs",'gear')
con <- c("drat","wt")
print(cat)
print(con)
```

### Question 5
### ggplot(mtcars, aes(x-disp, y=mpg)) This is an error: could not find function ggplot
### geom_point is missing

```{r}
library(ggplot2)

ggplot(mtcars, aes(x=disp, y=mpg)) + geom_point()
```

###Question 6
```{r}
data(mpg)
ggplot(data = mpg, aes(x=displ, y=hwy, color = factor(cyl))) + geom_point() + labs(color='Cylinders')
```
### leaving my comments here
### There seems to be a inverse/negative relationship between displ and hwy. As displ goes up hwy goes down.
### Each cylinder has varied hwy and displ

### Question 7
```{r}
ggplot(data = mpg, aes(x=displ, y=hwy, color = factor(cyl))) + geom_point() + facet_wrap(~drv) + labs(color='Cylinders')
```
### Rear wheel drives have more cylinders, have larger displ and lower hwy


